---
title: "Estimating the effects of Aducanumab on lifetime and ten-year risks of Alzheimer's disease"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Inflating treatment effects

Aducanumab has not been shown to have a massive treatment effect on the probability of developing Alzheimer's disease (AD). The relative risk (RR) for 1-year transitions is likely in the range of 0.75-1. However, it may be useful to see how a treatment with a larger effect would impact lifetime risks of AD. To that end, we assume RRs of 0.75, 0.50, and 0.25 to see how they change lifetime and ten-year risks.

# Lifetime Risks

## Females

We format tables such that each table represents a given starting state in the multistate model. Columns correspond to lifetime risks for each RR $\in (1, 0.75, 0.50, 0.25)$ for each age (rows). 

```{r}

library(knitr)
library(tidyverse)
library(here)

big <- readRDS(here("Aducanumab", "big.effects.rds"))

lifetime.f <- big$lifetime.f
lifetime.m <- big$lifetime.m
ten.f <- big$tenyear.f
ten.m <- big$tenyear.m

kable(lifetime.f$state1, caption = "Lifetime risks for females in state 1 (Normal)",
      format = 'simple')

kable(lifetime.f$state2, caption = 'Lifetime risks for females in state 2 (Amyloid)',
      format = 'simple')

```

\newpage

```{r}

kable(lifetime.f$state3, caption = 'Lifetime risks for females in state 3 (Neurodegeneration)',
      format = 'simple')

kable(lifetime.f$state4, caption = "Lifetime risks for females in state 4 (A + N)",
      format = 'simple')

kable(lifetime.f$state5, caption = "Lifetime risks for females in state 5 (A + N + MCI)",
      format = 'simple')
```

\newpage

## Males

```{r}
kable(lifetime.m$state1, caption = "Lifetime risks for males in state 1 (Normal)",
      format = 'simple')

kable(lifetime.m$state2, caption = 'Lifetime risks for males in state 2 (Amyloid)',
      format = 'simple')

kable(lifetime.m$state3, caption = 'Lifetime risks for males in state 3 (Neurodegeneration)',
      format = 'simple')

```

\newpage

```{r}

kable(lifetime.m$state4, caption = "Lifetime risks for males in state 4 (A + N)",
      format = 'simple')

kable(lifetime.m$state5, caption = "Lifetime risks for males in state 5 (A + N + MCI)",
      format = 'simple')
```

# Ten-year risks

## Females

```{r}

kable(ten.f$state1, caption = "Ten-year risks for females in state 1 (Normal)",
      format = 'simple')

```

\newpage

```{r}

kable(ten.f$state2, caption = "Ten-year risks for females in state 2 (Amyloid)",
      format = 'simple')

kable(ten.f$state3, caption = "Ten-year risks for females in state 3 (Neurodegeneration)",
      format = 'simple')

kable(ten.f$state4, caption = "Ten-year risks for females in state 4 (A + N)",
      format = 'simple')

kable(ten.f$state5, caption = "Ten-year risks for females in state 5 (A + N + MCI)",
      format = 'simple')

```

## Males

```{r}

kable(ten.f$state1, caption = "Ten-year risks for females in state 1 (Normal)",
      format = 'simple')

kable(ten.f$state2, caption = "Ten-year risks for females in state 2 (Amyloid)",
      format = 'simple')

kable(ten.f$state3, caption = "Ten-year risks for females in state 3 (Neurodegeneration)",
      format = 'simple')

```

\newpage

```{r}

kable(ten.f$state4, caption = "Ten-year risks for females in state 4 (A + N)",
      format = 'simple')

kable(ten.f$state5, caption = "Ten-year risks for females in state 5 (A + N + MCI)",
      format = 'simple')

```
